Online Multiclass Linear Classification with Bandit Feedback

### 2019-01-09 arxiv-version/algorithm-for-strongly-linearly-separable-data.tex
relwork, and some comments

>>>>

\chicheng{I think this part need to be formalized. Can we assume without loss of generality
that a deterministic prediction algorithm will have a smaller expected cost?
If so, define $F(k)$ be the number of mistake that will be made by the algorithm,
if the adversary commits to a label uniformly at random from $\cbr{1,2,\ldots,k}$,
and the guessing of the label lasts for $k$ rounds.
we can construct a recurrence of $F$: $F(k) = F(k-1) + \frac{k-1}{k}$, and $F(1) = 0$.
Therefore, $F(K) = \Omega(K)$.}

### 2019-01-09 arxiv-version/notions-of-linear-separability.tex
relwork, and some comments

\begin{definition}[Strong Linear separability]
Let $(V,\ip{\cdot}{\cdot})$ be an inner product space, $K$ be a positive
integer, and $\gamma$ be a positive real number. Let $(x_1, y_1), (x_2,
y_2), \dots, (x_T, y_T)$ be labeled examples in $V \times \{1,2,\dots,K\}$.

@@

The notion of linear separability with a margin is standard. It is used in the
full-information setting to upper bound the number of mistakes of the
\textsc{Multiclass Perceptron} algorithm. For completeness, we state the
algorithm in Appendix~\ref{section:multiclass-perceptron-proofs}. It is a
folklore result that if a set of labeled examples is separable with a margin
$\gamma$ and the norm of the examples is bounded by $R$ then \textsc{Multiclass
Perceptron} algorithm makes at most $\left\lfloor 2(R/\gamma)^2 \right \rfloor$
mistakes. Another folklore result is that \textsc{Multiclass Perceptron} is
essentially optimal in the sense that any deterministic algorithm must make
$\left\lfloor (R/\gamma)^2 \right \rfloor$ mistakes in the worst case. For
completeness, we give proofs of both of these results in
Appendix~\ref{section:multiclass-perceptron-proofs}.

The notion of strong linear separability is new.
\chicheng{I think the notion has appeared in~\cite{Chen-Chen-Zhang-Chen-Zhang-2009}, however,
they proposed an algorithm that doesn't achieve a finite mistake bound in the bandit
setting under this assumption (although they claimed that they have done so, the proof was
wrong..)}

It is easy to see that if labeled examples are strongly linearly
separable with margin $\gamma$ then they are linearly separable with margin
$2\gamma$. Indeed, if $w_1, w_2, \dots, w_K \in V$
satisfy \eqref{equation:strong-linear-separability-1},
\eqref{equation:strong-linear-separability-2},
\eqref{equation:strong-linear-separability-3} then they satisfy
\eqref{equation:linear-separability-1} and
\eqref{equation:linear-separability-2} with margin $\gamma' = 2\gamma$.

>>>>

\begin{definition}[Strong linear separability]
Let $(V,\ip{\cdot}{\cdot})$ be an inner product space, $K$ be a positive
integer, and $\gamma$ be a positive real number. Let $(x_1, y_1), (x_2,
y_2), \dots, (x_T, y_T)$ be labeled examples in $V \times \{1,2,\dots,K\}$.

@@

The notion of linear separability with a margin is standard. It is used in the
full-information setting to upper bound the number of mistakes of the
\textsc{Multiclass Perceptron} algorithm. For completeness, we state the
algorithm in Appendix~\ref{section:multiclass-perceptron-proofs}. It is a
folklore result that if a set of labeled examples is separable with a margin
$\gamma$ and the norm of the examples is bounded by $R$, then \textsc{Multiclass
Perceptron} algorithm makes at most $\left\lfloor 2(R/\gamma)^2 \right \rfloor$
mistakes. Another folklore result is that \textsc{Multiclass Perceptron} is
essentially optimal in the sense that any deterministic algorithm must make
$\left\lfloor (R/\gamma)^2 \right \rfloor$ mistakes in the worst case. For
completeness, we give proofs of both of these results in
Appendix~\ref{section:multiclass-perceptron-proofs}.

The notion of strong linear separability is new. Intuitively, strong linear separability
means that, for each $i$, the set of examples with class $i$ and the set of example of other classes are separated by a linear classifier $w_i$. Formally, $w_i$ separates the set of examples $\cbr{(x_t, y_t): t \in \cbr{1,\ldots,T}, y_t = i}$ and
$\cbr{(x_t, y_t): t \in \cbr{1,\ldots,T}, y_t \neq i}$.

\chicheng{I think the notion has appeared in~\cite{Chen-Chen-Zhang-Chen-Zhang-2009}, however,
they proposed an algorithm that doesn't achieve a finite mistake bound in the bandit
setting under this assumption (although they claimed that they have done so, the proof was
wrong..)}

It is easy to see that if a set of labeled examples are strongly linearly
separable with margin $\gamma$, then they are linearly separable with margin
$2\gamma$. Indeed, if $w_1, w_2, \dots, w_K \in V$
satisfy \eqref{equation:strong-linear-separability-1},
\eqref{equation:strong-linear-separability-2},
\eqref{equation:strong-linear-separability-3} then they satisfy
\eqref{equation:linear-separability-1} and
\eqref{equation:linear-separability-2} with margin $\gamma' = 2\gamma$.

### 2019-01-09 arxiv-version/related-work.tex
relwork, and some comments

\cite{Chen-Lin-Lu-2014, Zhang-Jung-Tewari-2018} study online bandit multiclass
boosting under bandit feedback, where one can view online boosting as online
linear classification by treating each base hypothesis as a separate feature.
\cite{Chen-Lin-Lu-2014}'s online weak learning condition implies that there is a
convex combination of base hypotheses that one-versus-all separates all examples
with a margin. Under this condition, it gives an algorithm with a $O(T^{3/4})$
mistake bound. \cite{Zhang-Jung-Tewari-2018} considers an online weak learning
condition that implies that there is a convex combination of base hypotheses
that weakly separates all examples with a margin
(See~\cite{Mukherjee-Schapire-2013}, Theorem 3). Under this condition, it gives
an algorithm with a $O(\sqrt{T})$ mistake bound with the knowledge of the edge
parameter. In addition, it gives an adaptive algorithm with a $O(T^{3/4})$
mistake bound.

The online bandit multiclass learning is a special case of the contextual bandits
problem~\cite{Auer-2003, Langford-Zhang-2007}, where the cost is the classification error.
A series of works in the contextual bandits learning literature focuses on
oracle-efficiency, that is, they assume access to an
oracle that given a set of cost-sensitve learning examples,

Recently, \cite{Foster-Krishnamurthy-2018} developed a rich theory on contextual bandits
with surrogate losses, focusing on benchmarks of the form $\min_{f \in \calF}
\frac{1}{K} \sum_{i=1}^K l_t(i) \phi( f_i(x) )$, where $\calF$ contains function $f$
such that $\sum_{i=1}^K f_i(\cdot) \equiv 0$, and $\phi(s) = \max(1 - \frac s
\gamma, 0)$ and $\min(1, \max(1 - \frac s \gamma, 0))$. On one hand, it gives
information-theoretic mistake upper bounds under various settings of $\calF$,
for instance parametric and nonparametric classes. On the other hand, it gives
an efficient algorithm that has a $O(\sqrt{T})$ regret against the benchmark of
$\calF = \cbr{x \mapsto W x: W \in \R^{k \times d}, \one^T W = 0}$.

>>>>

\cite{Chen-Lin-Lu-2014, Zhang-Jung-Tewari-2018} study online bandit multiclass
boosting under bandit feedback, where one can view online boosting as online
linear classification by treating each base hypothesis as a separate feature.
\cite{Chen-Lin-Lu-2014}'s online weak learning condition implies that
the set of examples is strongly separable by a convex combination of base hypotheses
with a margin. Under this condition, it gives an algorithm with a $O(T^{3/4})$
mistake bound. \cite{Zhang-Jung-Tewari-2018} considers an online weak learning
condition, which implies that all examples is separable by a convex combination of base hypotheses with a margin (See~\cite{Mukherjee-Schapire-2013}, Theorem 3).
Under this condition, it gives
a boosting algorithm with a $O(\sqrt{T})$ mistake bound with the knowledge of the edge
parameter. In addition, it gives an adaptive algorithm with a $O(T^{3/4})$
mistake bound.

There has been much literature on the contextual bandit learning problem~\citep{Auer-2003, Langford-Zhang-2007}.
In this problem, the learner is given a policy class $\Pi$, and
at each timestep $t$, an example $x_t$ is shown, with an associated cost vector $c_t$ hidden.
The learner then takes an action $\hat{y}_t$, and observed the incurred cost $c_t$.
The goal of the learner is to minimize its regret $\sum_{t=1}^T c_t(\hat{y}_t) - \min_{\pi \in \Pi} \sum_{t=1}^T c_t(\pi(x_t))$.
The online bandit multiclass learning problem is a special case of the contextual bandits problem,
where the cost $c_t(i)$ is the classification error $I(i \neq y_t)$,
and the policy class $\Pi$ is the set of linear classifiers
$\cbr{x \mapsto \argmax_y (Wx)_y: W \in \R^{k \times d}}$.
A series of works in the contextual bandits learning literature focus on
oracle-efficiency~\citep{DudikHKKLRZ11, Monster:2014, RakhlinS16, Syrgkanis2016a, Syrgkanis2016b}.
Specifically, they assume access to a policy optimization
oracle that receives a set of cost-sensitive learning examples and policy class $\Pi$,
return the policy in $\Pi$ that minimizes its empirical cost on the examples, and the algorithm
only call the oracle a polynomial number of times.
However, these algorithms are not truly computationally efficient in our setting,
as it is NP-hard in general to find a linear classifier that minimizes the empirical cost over a set of examples~\citep{arora1997hardness}.

%as algorithms generate cost-sensitive examples that may not be linearly separable
%Algorithms that satisfy oracle efficiency has been proposed in the literature, for
%example.

Recently, \cite{Foster-Krishnamurthy-2018} developed a rich theory of contextual bandits
with surrogate losses, focusing on benchmarks of the form $\min_{f \in \calF}
\sum_{t=1}^T \frac{1}{K} \sum_{i=1}^K c_t(i) \phi( f_i(x_t) )$,
where $\calF$ contains score functions $f$
such that $\sum_{i=1}^K f_i(\cdot) \equiv 0$, and $\phi(s) = \max(1 - \frac s
\gamma, 0)$ or $\min(1, \max(1 - \frac s \gamma, 0))$. On one hand, it gives
information-theoretic mistake upper bounds under various settings of $\calF$,
including parametric and nonparametric classes. On the other hand, it gives
an efficient algorithm that has a $O(\sqrt{T})$ regret against the benchmark of
$\calF = \cbr{x \mapsto W x: W \in \R^{k \times d}, \one^T W = 0}$. A direct
application of this result to the online bandit multiclass learning problem
yields an algorithm with a $O(\sqrt{T})$ mistake bound in the strong
linearly separable setting.
